%\section{FAQ}
\section{Bug Reporting and Enhancement Requests}

If you have access to the Sandia Software Bugzilla system, please use
the Bugzilla bug-tracking system. (Recommended) 

To get started with Bugzilla for the first time:
\begin{itemize}
\item Go to the Sandia Software Bugzilla page: 
  {\tt http://software.sandia.gov/bugzilla/}
\item Click on 'Open a new Bugzilla account'.
\item Enter your e-mail address and name.
\item Check your e-mail for your new password.
\item Sign in with your new password.
\item Configure your password and preferences how you want them. This
  can be done under 'Change password or user preferences' on the main
  page, or under 'Edit prefs' in the footer. 
\end{itemize}

To submit a bug, fill out a Bugzilla bug report:
\begin{itemize}
\item Go to the Sandia Software Bugzilla page:
  {\tt http://software.sandia.gov/bugzilla/}
\item Click 'Enter a new bug report' on the main page, or 'New' in the
  footer. 
\item Choose the package in which you want to request a feature or
  report a bug/issue (or Trilinos if it applies to the entire Trilinos
  framework). 
\item Read the 'bug writing guidelines,' check the list of 'most
  frequently reported bugs,' and 'search' for the bug to see if the
  bug has already been submitted (in which case, it may be helpful to
  add a new comment to the existing report). 
\item Select the version number, component, platform, operating
  system, priority, and severity of the bug. 
\item Enter a concise, informative summary. 
\item Describe, in as much detail as possible, the feature you are
  requesting or the bug/issue you are reporting. 
\end{itemize}

To find a bug, run a Bugzilla query: 
\begin{itemize}
\item Go to the Sandia Software Bugzilla page:
  {\tt http://software.sandia.gov/bugzilla/}
\item Click 'Query existing bug reports' on the main page, or 'Query'
  in the footer. 
\item Select appropriate entries from any relevant fields.
\item Click the 'Search' button.
\end{itemize}


If you do not have access to the Sandia Software Bugzilla system,
please send an e-mail describing the bug in as much detail as possible 
to vehowle@software.sandia.gov.  

\section{Mailing Lists}
Meros maintains a number of email lists that users and developers can
subscribe to. We recommend subscribing to the following mailing lists
as appropriate. 

\begin{itemize}
\item Meros-Announce Mailing List - Low-volume, primarily for Meros
  release announcements. (Highly Recommended) 
\item Meros-Users Mailing List - Discussion forum for users of
  Meros. (Recommended) 
\item Meros-Developers Mailing List - Discussion forum for Meros
  developers. (Developers) 
\item Meros-Checkins Mailing List - CVS Commit Message. (Active
  Developers) 
\item Meros-Regression Mailing List - Captures output from Meros
  automated regression tests. (Active Developers) 
\end{itemize}




