\section{Getting Started}

\subsection{Downloading Meros and Trilinos}
Meros is a package within the Trilinos Project and currently can be
downloaded and built with Trilinos.
The Trilinos software, including Meros, can be downloaded from
\begin{itemize}
\item[]
  {\tt http://trilinos.sandia.gov/}. 
\end{itemize}
Follow the download instructions on
the web site.



\subsection{Licensing}
Meros is available for download (through Trilinos) under the GNU
Lesser General Public License~\cite{gnuLGPL}.


\subsection{Configuration, Compilation, and Installation}
\begin{itemize}
\item System and software requirements
\item Configuring and building Trilinos in general
\item Configuring and building Meros within Trilinos
\item Testing your build
\end{itemize}



Using the instructions here, your build of trilinos will have the following
libraries: aztecoo, epetra, epetraext, ML, newpackage,
nox teuchos, and thyra.
\begin{itemize}
\item \verb!cd! into the trilinos~directory.
\item Make a build directory, e.g., \verb!mkdir LINUX!.
\item \verb!cd LINUX!.
\item Configure trilinos:
  \begin{enumerate}
  \item   If you do not want to use MPI:
\begin{verbatim} 
{ ../configure --enable-teuchos --enable-amesos} 
\end{verbatim}
  \item To use MPI:
\begin{verbatim} 
{ ../configure  --enable-teuchos \\
  --enable-amesos --with-mpi-compilers=/usr/local/mpich/bin} 
\end{verbatim}
where your path to the mpi-compilers is specified.
  \end{enumerate}

\item Build trilinos: \verb!make!.
\item If your build finished without errors, you should see the directory\\
\verb!Trilinos/LINUX/packages/!, with subdirectories below that for
each individual library.   Meros's subdirectory, \verb!meros!, should contain
files \verb!config.log!, \verb!config.status!, \verb!Makefile!, and
\verb!Makefile.export!, and directories \verb!src! and \verb!examples!.
Directory \verb!src! contains object files and \verb!libmeros.a!.
Directory \verb!examples! contains executables with extension \verb!.exe!,
symbolic links to the corresponding source code, and object files.
Directory \verb!test! is intended primarily for developers and can be ignored.
\item Look in \verb!Trilinos/LINUX/packages/meros/examples! for
examples of how to use Meros. File
\verb!Trilinos/packages/meros/examples/README! suggests how to use the
examples.
\end{itemize}


Sample configure script for Meros:

\begin{verbatim}
BUILD_DIR=`pwd`

../configure \
--enable-mpi \
--with-mpi-compilers="/usr/local/mpich-1.2.7_gcc3.4.3/bin" \
CXXFLAGS="-g -O0 -Wreturn-type" \
--with-libs="-lexpat" \
--disable-default-packages \
--enable-teuchos --enable-teuchos-extended \
--enable-teuchos-complex --enable-teuchos-abc --enable-teuchos-expat \
--enable-thyra \
--enable-thyra-examples \
--enable-epetra \
--enable-epetra-thyra \
--enable-epetraext \
--enable-epetraext-thyra \
--enable-ml \
--enable-ml-thyra \
--enable-aztecoo \
--enable-aztecoo-thyra \
--enable-ifpack \
--enable-ifpack-thyra \
--enable-nox \
--enable-stratimikos \
--enable-meros \
--enable-meros-examples \
--with-gnumake \
--with-install="/usr/bin/install -c -p" \
--prefix=$BUILD_DIR
\end{verbatim}

